---
title: "Lab 5"
author: "Jonathan Kanyinda"
date: "2021-02-23"

output: github_document 
---

Welcome to my Lab 5 report! In this lab report I will use the data from Edgar Anderson's Iris data to rename variables, create new variables, summarize a dataset, summarize a dataset by groups and visualize the relationship between variables (Numerical and categorical, two numerical variables and three variables).

## Load packages

```{r}
library(tidyverse) # load tidyverse package
```

## Read data

```{r}
iris
iris <- as_tibble(iris)
iris

```

## Question 1

```{r}

# Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table

question1_table <- rename (iris, sepal_length = Sepal.Length, sepal_width = Sepal.Width, petal_length = Petal.Length, petal_width = Petal.Width, species = Species) 
question1_table 
```

## Question 2

```{r}

# Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.

question2_table <- mutate(question1_table,sepal_length_cm = sepal_length/10, sepal_width_cm = sepal_width/10, petal_length_cm = petal_length/10, petal_width_cm = petal_width/10 ) 

question2_table 
```

## Question 3 

```{r}
# Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.

t3 <- mutate(question2_table, sepal_area = sepal_length * sepal_width,petal_area = petal_length * petal_width ) 

question3_table <- select(t3, sepal_area, petal_area, species) 

question3_table  

```

## Question 4

```{r}


#Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table: a)sample size. b)maximum value. c) minimum value. d)range.e)median. f) first quartile (q1). g)third quartile (q2). h)inter-quartile range (iqr)

t4 <- select(question1_table, sepal_length) 

question4_table <-summarize( t4,sampl_size = n(),
    max_value = max(sepal_length), 
    min_value = min(sepal_length), 
    range = max_value - min_value,
    median = median(sepal_length), 
    q1 = quantile(sepal_length, probs = 0.25), 
    q2 = quantile(sepal_length, probs = 0.75), 
    IQR =  IQR(sepal_length) 
  )

question4_table 



```

## Question 5 

```{r}
# Calculate the following statistics for each species from the petal width variable and print the resulting table: a) sample size. b)mean.  c)standard deviation.  d)variance.  e)standard error of the mean. f)approximate 95% confidence interval.

iris_grouped <- question1_table %>% group_by(species)%>%
  summarize(
    sample_size = n(),
    mean = mean(petal_width),
    standar_deviation = sd(petal_width),
    variance = var(petal_width), 
    sem = standar_deviation/sqrt(sample_size),
    ci_upper_limit = mean + 2 * sem,
    ci_lower_limit = mean - 2 * sem
  )

iris_grouped

```

## Question 6

```{r}
# Visualize the relationship between petal width and species using a strip plot.

tab_6 <- select(question1_table, species , petal_width)

tab_6

ggplot(data = question1_table) +
  geom_jitter(mapping = aes(x = species, y = petal_width))


```

## Question 7

```{r}
# Starting with the previous graph, add the mean and 95% confidence interval for each species

ggplot(data =question1_table) +
  geom_jitter(mapping = aes(x = species, y = petal_width)) +
  geom_crossbar(
    data = iris_grouped, 
    mapping = aes(x = species, y = mean, ymax = ci_upper_limit, 
                  ymin = ci_lower_limit),
    color = "red"
  )

```

## Question 8

```{r}
# Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.

ggplot(data = question1_table ) +
  geom_point(mapping = aes(x = petal_length, y =petal_width , color = species))

```

## Session info

```{r session-info}
sessioninfo::session_info()
```









